%% example.tex
%% Jeremy Singer
%% 16 Oct 12

\documentclass{mpaper}

\begin{document}

\title{IPython Observatory}
\author{Silviya Sotirova}
\matricnum{1103571s}

\maketitle

\begin{abstract}

Scientists have been adopting various new tools so that they can maintain, share and replace long data science workflows. One of the most popular approaches is the usage of software code notebooks, which gain most of its attractiveness with the support of IPython Notebook. Since 2001, IPython has grown out of the scientific computing community and it has developed features that are useful for data scientists. IPython has been developed for powerful and practical interaction with data. It also addresses other important aspects for scientists, such as reproducibility and collaboration - issues that are equally important to data scientists working in industry. However, little is known about how iPython has affected working practices for scientific programmers.

The project is analyzing in detail the IPython Notebook usage into the fields of scientific research and software development. A popular and effective place for sharing various researches and contributions for projects is the open source hosting service GitHub. This research is investigating GitHub, since it allows tracking of data changes, commits, issues and documentation. 

\end{abstract}
\section{Introduction}

\input{Introduction/intro}

\section{Background}
\input{background/background.tex}

\section{Analysis}
\input{analysis/analysis.tex}

\section{Discussion \& Conclusions}
\input{conclusions/conclusions.tex}

\vspace{5mm}
\bibliographystyle{abbrv}
\bibliography{bibliography}


\end{document}